Measurement Clusterization in Bayesian D-optimal Designs in Infinite Dimensions

### 2024-07-21 latex/ack.tex
before diff

This study is a result of research I started during my PhD studies
under the instruction of Prof. Georg Stadler at the Courant Institute
of Mathematical Sciences. I would like to thank him for his great
mentorship, attention to details and kindness. I would also like to
thank Christian Remling, who helped me find a proof for Lemma
\ref{lemma:free} in
\href{https://mathoverflow.net/questions/280168/redistribute-diagonal-entries-of-a-matrix/280203#280203c}{Mathoverflow}.

>>>>

This study is a result of research I started during my PhD studies
under the instruction of Prof.~Georg Stadler at the Courant Institute
of Mathematical Sciences. I would like to thank him for his great
mentorship, attention to details and kindness. I would also like to
thank Christian Remling, who helped me find a proof for Lemma
\ref{lemma:free} in
\href{https://mathoverflow.net/questions/280168/redistribute-diagonal-entries-of-a-matrix/280203#280203c}{Mathoverflow}.
Last, but certainly not least, I would like to thank the three
referees, associate editor and editor in chief for providing detailed
and insightful reviewes that have made this manuscript a whole lot
better.
%Having worked on this manuscript alone for several years, they have
%almost become my coautoh

### 2024-07-21 latex/discussion.tex
before diff

\subsubsection{Answer to Question \ref{q:why}}
Armed with Theorem \ref{thm:char} we can give a compelling explanation
of the measurement clusterization we obeserved in
Fig.~\ref{fig:}. From part (..) of the Theorem \ref{thm:char} and
Fig.~\ref{fig:tikz clusterization}, we know that a D-optimal design
aims to reduce uncertainty of eigenvectors of
$(\fwd\prcov\fwd^*)^{-1}$ where prior covariance eigenvalues
$\lambda_j$ are highest. D-optimal design will try to focus on the
leading eigenvectors of $(\fwd\prcov\fwd^*)^{-1}$ and take
measurements where other \emph{scaled} eigenvectors are close to
zero.

For concreteness, let us assume that in a certain problem, a D-optimal
design aims to measure only the two leading eigenvectors, i.e.~$\rank
\obs^*\obs =: k=2$. Thus, a D-optimal design will (try to) ignore all
eigenvectors corresponding to $j > 2$. Since $\lambda_j$ decay to
zero, a D-optimal design should only try to avoid eigenvectors for
which $\lambda_j$ is not already small --- e.g.~eigenvectors
$j=3,4$. It is possible that the geometry of the domain $\Omega$ and
the structure of the eigenvectors leave only a few locations in
$\Omega$ where the amplitude of eigenvectors $j=1,2$ is high while the
amplitude of eigenvectors $j=3,4$ is low. In such case, clusterization
will occur whenever the number of measurements $m$ is large enough, by
the pigeonhole principle. This scenario is illustrated for the inverse
problem of the 1D heat equation in Fig.~\ref{fig:eigenvectors}.

\begin{figure}\label{fig:eigenvectors}
    \centering
    \includegraphics[width=\textwidth]{eigenvectors_dst.pdf}
    \caption{D-optimal measurement locations ($m=4$ measurements) and
      weighted eigenvectors for finding the initial condition of
      the 1D heat equation. Measurement locations and weighted
      eigenvectors are plotted over the computational domain $\Omega =
      [0, 1]$ (x-axis). Measurement clusterization occurs
      approximately at $0.31$ and $0.69$. These two locations are a
      compromise between zeros of eigenvectors a D-optimal design aims
      to ignore (third and up) and staying far from zeros of the
      eigenvectors a D-optimal design aims to measure (first and
      second). Allocating $m=4$ measurements into two locations
      results in clusterization, according to the pigeonhole
      principle.}
  \label{fig:why}
\end{figure}

By Weyl's law, eigenvalues $\lambda_j$ of the Laplace operator
increase as $\lambda_j \sim j^{d/2}$, where $d = \dim \Omega$ is the
dimensionality of the domain. In accordance with Weyl's law,
\cite[Theorem 3.1]{Stuart10} suggests taking priors as
"Laplacian-like" operators to power smaller than $-\frac{d}{2}$, to
maintain regularity of posterior realizations. Thus, even though the
computational domain is getting larger and sparser as the dimension
$d$ grows, clusterization is not necessarily harder to achieve.

%% In our model we did not specify any specific prior, so it is possible
%% that one will be able to devise a prior that will not result in
%% clusterization. Indeed, if the rate of increase of prior precision
%% eigenvalues is small,

>>>>

\subsubsection{Answer to Question \ref{q:why}}
Theorem \ref{thm:char} helps us give a compelling explanation of the
measurement clusterization we obeserved for the inverse problem of the
heat equation (Fig.~\ref{fig:}). Below, we take the following steps:
We first utilize Theorem \ref{thm:char} to understand what a D-optimal
design will look like for our relaxed model; we then argue that
although not identical, a D-optimal design for the full inverse
problem of the heat equation shares similar charachteristics to the
relaxed design; finally, we show how ...

First, recall from Theorem \ref{thm:char} that in our relaxed model,
for a D-optimal design $\opt$:
\begin{enumerate}
  \item $\opt^*\opt$ is simultaneously diagonalizable with $\fwd
    \prcov \fwd^*$.
  \item $\opt^*\opt$ has non-zero eigenvalues only for the $k:=\rank
    \opt^*\opt$ smallest eigenvalues $\lambda_j$ of
    $\fwd\prcov\fwd^*$.
  \item $k$ is monotonic in $m$, the number of measurements we take.
\end{enumerate}

Now, consider $\fwd$ and $\prcov$ from the \emph{the inverse problem
of the heat equation}. As before, we denote the eigenvalues of
$\fwd\prcov\fwd^*$ by $\lambda_j$. We input these eigenvalues into our
\emph{relaxed} model, and find a D-optimal design $\opt$. In our
relaxed model, the measurements we take are best utilized in reducing
uncertainty for the first $k$ eigenvectors. So, a D-optimal design
arising from our \emph{relaxed model} completely avoids measuring
eigenvectors $k+1$ and above.

Of course, in a real life problem --- such as the inverse problem of
the heat equatiion --- it is likely impossible to find measurement
locations for which all eigenvectors $k+1$ and above are
zero. However, we can still expect a D-optimal design to give as
little weight as possible to eigenvectors $k+1$ and above. Thus a
D-optimal design is a trade-off between measuring eigenvectors we want
(eigenvectors for which $\lambda_j$ is large) and treating the rest as
noise.

We explore the abovementione trade-off for the inverse problem of the
heat equation in Fig.~\ref{fig:eigs}. We allow $m=4$ measurements in
$[0,1]$ and observe that D-optimal measurement locations are clustered
at $x_1 = 0.31$ and $x_2 = 0.69$. Upon close inspection of the scaled
eigenvectors of $\fwd \prcov \fwd^*$, we first observe that
eigenvectors $3$ and above have negligible prior amplitude. Since we
only have $m=4$ measurements at our disposal, Theorem \ref{thm:char}
tells us we should only care about measuring the first and second
eigenvectors. Then, we note the D-optimal $x_1,x_2$ present a
compromise between the amplitude of the first and second
eigenvectors. For example, a measurement at $x=0.5$ would have ignored
the second eigenvector altogether, since its value is zero at $x=0.5$.
Now we can understand measurement clusterization: attempting to
measure what we want and ignoring the rest, along with the spatial
limitations on measurements imposed by the structure of eigenvectors
of $\fwd \prcov \fwd^*$ is the cause for measurement clusterization in
D-optimal designs.

%% In the inverse problem of the heat equation, \ev_3, \ev_4,\dots$
%% implies we seek measurement locations for which $\lambda_1\ev_1,
%% \lambda_2\ev_2$ are large, while $\lambda_3\ev_3,
%% \lambda_3\ev_3,\dots$ are small. It turns out that there are two
%% locations that give the best trade-off, and measurements are taken in
%% those two locations. Once we have more than two measurements at our
%% disposal, measurement locations becomde clustered, by the pigeonhole
%% principle, see Fig.~\ref{fig:eigenvectors}.

%% The way a D-optimal design tries to ignore all eigenvectors
%% corresponding to $j > 2$. Since $\lambda_j$ decay to zero, a D-optimal
%% design should only try to avoid eigenvectors for which $\lambda_j$ is
%% not already small --- e.g.~eigenvectors $j=3,4$. It is possible that
%% the geometry of the domain $\Omega$ and the structure of the
%% eigenvectors leave only a few locations in $\Omega$ where the
%% amplitude of eigenvectors $j=1,2$ is high while the amplitude of
%% eigenvectors $j=3,4$ is low. In such case, clusterization will occur
%% whenever the number of measurements $m$ is large enough, by the
%% pigeonhole principle. This scenario is illustrated for the inverse
%% problem of the 1D heat equation in Fig.~\ref{fig:eigenvectors}.

\begin{figure}\label{fig:eigenvectors}
    \centering
    \includegraphics[width=\textwidth]{eigenvectors_dst_scaled.png}
    \caption{D-optimal measurement locations ($m=4$ measurements) and
      weighted eigenvectors for finding the initial condition of
      the 1D heat equation. Measurement locations and weighted
      eigenvectors are plotted over the computational domain $\Omega =
      [0, 1]$ (x-axis). Measurement clusterization occurs
      approximately at $0.31$ and $0.69$. These two locations are a
      compromise between zeros of eigenvectors a D-optimal design aims
      to ignore (third and up) and staying far from zeros of the
      eigenvectors a D-optimal design aims to measure (first and
      second). Allocating $m=4$ measurements into two locations
      results in clusterization, according to the pigeonhole
      principle.}
  \label{fig:why}
\end{figure}

### 2024-07-21 latex/intro.tex
before diff

\section{Introduction}\label{section:intro}
The practice of experimental design entails specifying every detail of
running an experiment \cite{chaloner}: at its simplest form,
experimental design involves choosing a set of measurements to be
taken. In a medical experiment, experimental design may involve
choosing participants to whom some drug will be administered, and
measuring their blood pressure. Experimental design also arises 
process of searching for oil. Since measuremnts are expensive and
require digging holes in the ground, only a limited number of
measuremnts can be taken, hence carefully choosing measuremt locations
is crucial.

Specifiyng the right set of measurements holds particular significance
when solving an \emph{inverse problem} --- i.e. making inference of
the physical world utilizing a physical model of some phenomenon of
interest. In an inverse problem, we seek to infer physical quantities,
based on observations and a model of the physical world, typically
phrased in the language of ordinary or partial differential equations.

Inverse problems are ubiquitous in science and engineering, and we
will now give several illustrative examples. In electric impedance
tomography (EIT) we seek to infer the impedance of the inside of a
human body from observations of electrical current impedence measured
at specific electrode locations on the skin, leveraging known
equations of electric current flow
\cite{horesh2010impedance}. Magnetic Resonance Imaging (MRI) also
entails solving an inverse problem, where radio-frequency pulses are
sent through the human body in the presence of a strong magnetic field
and the response of our bodies' contents to such perutrbations is
measured, allowing a radiologist to view the internals of a body in a
noninvasive manner \cite{horesh2008mri}. Yet another example of an
inverse problem arises in oil and gas exploration, where acoustic
waves are sent through "boreholes" --- deep cylindrical wells drilled
into the ground --- and data of travel and return times of said waves
are recorded. Since travel and return time is influenced by the
properties of the subsurface materials, such as density, elasticity,
and the presence of fluids or voids, we can combine travel data with a
geophisical models of contents of the earth's crust to facilitate
reconstructing and inferring the structure of the subsurface in a
process called \emph{borehole tomography}
\cite{horesh2008borehole}. Inverse problems also arise in many other
areas of seismology and geology \cite{rabinowitz1990, steinberg1995}
and medical imaging \cite{tarantola}. While in many inverse problems
the goal is to infer some parameter, in all of the above examples and
in the formulation we consider here, the goal is to infer some
\emph{function} over $ \Omega$, where \(\Omega \subseteq \mathbb{R}^d,
d=1,2,3\) is the spatial domain of interest.

@@

For concreteness, let us consider a toy model of an inverse problem:
the \emph{heat equation} in one dimension (1D heat equation
henceforth, see Section \ref{subsec:1d heat} for more details). The
heat equation is a partial differential equation which is in fact a
set of three equations: eqs.~\eqref{eq:heat1}, \eqref{eq:heat2}, and
\eqref{eq:heat3}. Readers less familiar with partial differential
equations should think of the heat equation as follows: before us
there is a perfectly insulated metal rod. At each point on the rod the
temperature is different and unknown, and we call this temperature
distribution the \emph{initial condition}. We wait for a short time
and let the heat dissipate a bit inside the rod. The final heat
distribution in the rod is dictated by the following set of equations
which defines the 1D heat equation:
\begin{subequations}%\label{eq:heat equation}
  \begin{alignat}{2}
    u_t &= \alpha \Delta u &&\qquad \text{in } [0,1] \times [0,\infty), \label{eq:heat1}\\
    u &= 0 &&\qquad \text{on } \{0, 1\} \times [0,\infty), \label{eq:heat2}\\
    u &= u_0 &&\qquad \text{on }[0,1] \times \{0\}. \label{eq:heat3}\\
  \end{alignat}
\end{subequations}

As time passes, heat disspates across the rod as hotter regions,
e.g.~local maximas of the heat distribution, become cooler. This
behavior is captured by eq.~\eqref{eq:heat1}: the Laplacian at local
maxima is negative, so $u_t = \Delta u$ implies $u$ decreases at its
maximas, and the reverse happens at local minima. Eq.~\eqref{eq:heat2}
describes the \emph{boundary condition}, i.e.~the interaction of the
heat inside the rod with the boundary of the domain $\Omega =
[0,1]$. Here we consider a \emph{homogeneous Dirichlet boundary
condition} which is best understood as an absorbing boundary: heat
that interacts with the boundary immediately
disappears. Eq.~\eqref{eq:heat3} describes the \emph{initial
condition} of the heat, i.e.~the heat distribution at time $t=0$.

So far, we described the \emph{forward} problem: how heat evolves in
time. In the inverse problem, our goal is to infer the initial
condition $u_0 = u(\cdot, 0)$ from observations of the final condition
$u(\cdot, T)$. We are allowed to take some number of temperature
measurements on the rod once our waiting time has passed. The inverse
problem of the 1D heat equation is difficult ("ill-posed") since heat
spreads in a diffusive manner and any roughness in the initial
condition is quickly smoothed, see Supplementary movies S1 and S2. We
will give a fully Bayesian formulation of a generic linear inverse
problem in Section \ref{section:prelims}.

@@

%% first: blue, second: red, third: green, fourth: black,
%% fifth:magenta
\begin{figure}
    \centering
    \includegraphics[height=0.5\textwidth]{dst_modelError0.pdf}
    \caption{Measurement clusterization for D-optimal designs for the
      inverse problem of the heat equation. Measurement locations were
      chosen according to the Bayesian D-optimality criterion of
      Theorem \ref{thm:d optimality}. Measurement locations are
      plotted over the computational domain \(\Omega = [0, 1]\)
      (x-axis), for varying numbers of measurements (y-axis). The
      colored numbers are measurement indices, plotted for visual
      clarity. Measurement clusterization already occurs for three
      measurements: the second measurement (red) is overlaid on the
      third (green). For five measurements, first (blue) and second
      (red) measurements are clustered, as well as the fourth (black)
      and the fifth (magenta).}
  \label{fig:clusterization illustration}
\end{figure}

Clusterization should not be confused with either repetition nor with
replication, which are commonly viewed as beneficial and even
necessary aspects of an optimal design \cite{morris, fisher
  schafer2001replication}. For example, in his famous milk and tea
experiment, \cite[Section 11]{fisher} suggested that repetition is a
"way of enlarging the experiment and, thereby, increasing its
sensitiveness". In another example, \cite{fay2000rainfall} measured
the effect of rainfall on grass growth in plots of land. The
experiment involved fifteen "rainfall manipulation shelters", where
"Four rainfall manipulation treatments (three replicates) then were
assigned to 12 of the plots". It seems reasonable the researchers
replicate the phenomenon they are trying to study. Clusterization
seems like a whole different beast: a clustered design in the rainfall
experiment would imply the researchers should take a repeated
measurement \emph{on the same plot}, at the expense of measuring grass
growth in other plots. In sharp contrast to repetition and
replication, clusterization is highly nonintuitive and, as we will
see, its origins run considerably deeper.

@@

\subsection{Contribution}
The primary objective of this study is to provide a comprehensive
understanding of measurement clusterization by addressing the
aforementioned questions. Our focus centers around investigating the
Bayesian D-optimality criterion. We conduct an analysis of Bayesian
D-optimal designs within the context of linear inverse problems over
Hilbert spaces. We propose a novel relaxed model for D-optimality that
maintains analytical tractability and enables the identification of
D-optimal designs using Lagrange multipliers. This analytical
framework facilitates the exploration of the questions posed at the
end of the previous paragraph:

@@

The implication of the answers we give to Questions \ref{q:generic},
\ref{q:mitigate}, and \ref{q:why} is two-fold:

The implication of our answer to Question \ref{q:generic} is that
encountering clusterization should be expected in many different
inverse problems across many different scientific fields and
disciplines. Thus, researchers that encounter clusterization should
not be surprised or wary. Then, in our answer to Question \ref{q:why},
we explain what is the source of clusterization. As it appears, the
source is universal: a D-optimal design reduces uncertainty for a
select set of prior covariance eigenvectors --- those with the most
prior uncertainty, i.e.~those the practitioner cares about the most!
Consequently, practitioners should not try to avoid measurement
clusterization. Rather, practitioners should should take repeated
measurements (e.g.~in MRI and borehole tomography), increase apparatus
sensitivity (e.g.~in EIT), or take consecutive measurements (e.g.~in
the 1D heat equation). Overall, we show that measurement
clusterization is a natural and (almost) inevitable part of Bayesian
D-optimal designs.

>>>>

\section{Introduction}\label{section:intro}
The practice of experimental design entails specifying every detail of
running an experiment \cite{chaloner}: at its simplest form,
experimental design involves choosing a set of measurements to be
taken. For example, in a medical experiment, experimental design may
involve choosing participants to whom some drug will be administered,
and measuring their blood pressure. Experimental design also arises in
the process of searching for oil. Since measuremnts are expensive and
require digging holes in the ground, only a limited number of
measuremnts can be taken, hence carefully choosing measuremt locations
is crucial.

Specifiyng the right set of measurements holds particular significance
when solving an \emph{inverse problem} --- i.e. making inference of
the physical world utilizing a physical model a phenomenon of interest
\cite{tarantola,somersalo}. In an inverse problem, we seek to infer
physical quantities, based on observations and a model of the physical
world. Models are typically phrased in the language of ordinary or
partial differential equations.

Inverse problems are ubiquitous in science and engineering, and we
will now give several illustrative examples. In electrical impedance
tomography (EIT) we seek to infer the impedance of the inside of a
human body from observations of electrical current impedance measured
at specific electrode locations on the skin, leveraging known
equations of electric current flow
\cite{horesh2010impedance}. Magnetic Resonance Imaging (MRI) also
entails solving an inverse problem, where radio-frequency pulses are
sent through the human body in the presence of a strong magnetic field
and the response of one's body contents to such perutrbations is
measured, allowing a radiologist to view the internals of a body in a
noninvasive manner \cite{horesh2008mri}. Yet another example of an
inverse problem arises in oil and gas exploration, where acoustic
waves are sent through "boreholes" --- deep cylindrical wells drilled
into the ground --- and data of travel and return times of said waves
are recorded. Travel and return time is influenced by the properties
of the subsurface materials, such as density, elasticity, and the
presence of fluids or voids, combining travel data with a geophisical
models of contents of the earth's crust to facilitate reconstructing
and inferring the structure of the subsurface in a process called
\emph{borehole tomography} \cite{horesh2008borehole}. Inverse problems
also arise in many other areas of seismology and geology
\cite{rabinowitz1990, steinberg1995} and medical imaging
\cite{tarantola}. While in many inverse problems the goal is to infer
some parameter, in all of the above examples and in the formulation we
consider here, the goal is to infer some \emph{function} over $
\Omega$, where (usually) \(\Omega \subseteq \mathbb{R}^d, d=1,2,3\) is
the spatial domain of interest.

@@

For concreteness, let us consider a toy model of an inverse problem:
the \emph{heat equation} in one dimension (1D heat equation
henceforth, see Section \ref{subsec:1d heat} for more details). The
heat equation is a partial differential equation compose of three
equations: eqs.~\eqref{eq:heat1}, \eqref{eq:heat2}, and
\eqref{eq:heat3}. Readers less familiar with partial differential
equations should think of the heat equation as follows: before us
there is a perfectly insulated metal rod. At each point on the rod the
temperature is different and unknown, and we call this temperature
distribution the \emph{initial condition}. We wait for a short time
$T$ and let the heat dissipate a bit inside the rod. The final heat
distribution in the rod is dictated by the following set of equations
which defines the 1D heat equation:
\begin{subequations}%\label{eq:heat equation}
  \begin{alignat}{2}
    u_t &= \alpha \Delta u &&\qquad \text{in } [0,1] \times [0,\infty), \label{eq:heat1}\\
    u &= 0 &&\qquad \text{on } \{0, 1\} \times [0,\infty), \label{eq:heat2}\\
    u &= u_0 &&\qquad \text{on }[0,1] \times \{0\}. \label{eq:heat3}\\
  \end{alignat}
\end{subequations}

As time passes, heat disspates across the rod as hotter regions,
e.g.~local maximas of the heat distribution, become cooler. This
behavior is captured by eq.~\eqref{eq:heat1}: the Laplacian at local
maxima is negative, so $u_t = \Delta u$ implies $u$ decreases at its
maximas, and the reverse happens at local minima. Eq.~\eqref{eq:heat2}
describes the \emph{boundary condition}, i.e.~the interaction of the
heat inside the rod with the boundary of the domain $\Omega =
[0,1]$. Here we consider a \emph{homogeneous Dirichlet boundary
condition} which is best understood as an absorbing boundary: heat
that interacts with the boundary is immediately
absorbed. Eq.~\eqref{eq:heat3} describes the \emph{initial condition}
of the heat, i.e.~the heat distribution at time $t=0$. Thus, we have
described the \emph{forward} problem: how heat evolves in time,
starting at $t=0$ up until $t=T$.

In the inverse problem of the 1D heat equation, our goal is to infer
the initial condition $u_0 = u(\cdot, 0)$ from observations of the
final condition $u(\cdot, T)$. We are not able to measure $u(\cdot,
T)$ at every point in the domain $\Omega$. Rather, we take some number
of temperature measurements on the rod once our waiting time $T$ has
passed and try to infer $u(\cdot, 0)$ from these measurements.

The inverse problem of the 1D heat equation is difficult ("ill-posed")
since heat spreads in a diffusive manner and any roughness in the
initial condition is quickly smoothed, see Supplementary movies S1 and
S2. For ill-posed problems regularization is in order, and we
implement such regularization via a fully Bayesian formulation of the
inverse problem (see Section \ref{section:prelims} for details).

@@

%% first: blue, second: red, third: green, fourth: black,
%% fifth:magenta
\begin{figure}
    \centering
    \includegraphics[height=0.5\textwidth]{dst_modelError0.png}
    \caption{Measurement clusterization for D-optimal designs for the
      inverse problem of the heat equation. Measurement locations were
      chosen according to the Bayesian D-optimality criterion of
      Theorem \ref{thm:d optimality}. Measurement locations are
      plotted over the computational domain \(\Omega = [0, 1]\)
      (x-axis), for varying numbers of measurements (y-axis). The
      colored numbers are measurement indices, plotted for visual
      clarity. Measurement clusterization already occurs for three
      measurements: the second measurement (red) is overlaid on the
      third (green). For five measurements, first (blue) and second
      (red) measurements are clustered, as well as the fourth (black)
      and the fifth (magenta).}
  \label{fig:clusterization illustration}
\end{figure}

Clusterization should not be confused with either repetition nor with
replication, which are commonly viewed as beneficial and even
necessary aspects of an optimal design \cite{morris, fisher
  schafer2001replication}. For example, in his famous milk and tea
experiment, \cite[Section 11]{fisher} suggested that repetition is a
"way of enlarging the experiment and, thereby, increasing its
sensitiveness". In another example, \cite{fay2000rainfall} measured
the effect of rainfall on grass growth in plots of land. The
experiment involved fifteen "rainfall manipulation shelters", where
"Four rainfall manipulation treatments (three replicates) then were
assigned to 12 of the plots". It seems reasonable the researchers
replicate the phenomenon they are trying to study. Clusterization is
different: a clustered design in the rainfall experiment would imply
the researchers should take a repeated measurement \emph{on the same
plot}, at the expense of measuring grass growth in other plots. In
sharp contrast to repetition and replication, clusterization is highly
nonintuitive and, as we will see, its origins run considerably deeper.

@@

\subsection{Contribution}
The primary objective of this study is to provide a deep understanding
of measurement clusterization by addressing the aforementioned
questions. Our focus centers around investigating the Bayesian
D-optimality criterion. We conduct an analysis of Bayesian D-optimal
designs within the context of linear inverse problems over Hilbert
spaces. We propose a novel relaxed model for D-optimality that
maintains analytical tractability and enables the identification of
D-optimal designs using Lagrange multipliers. This analytical
framework facilitates the exploration of the questions posed at the
end of the previous paragraph:

@@

The implication of our answer to Question \ref{q:generic} is that
encountering clusterization should be expected in many different
inverse problems across many different scientific fields and
disciplines. Thus, researchers that encounter clusterization should
not be surprised or wary. Then, in our answer to Question \ref{q:why},
we explain what is the cause of clusterization. As it appears, the
cause is universal: a D-optimal design reduces uncertainty for a
select set of prior covariance eigenvectors --- those with the most
prior uncertainty, i.e.~those the practitioner cares about the most!
Consequently, practitioners should not try to avoid measurement
clusterization. Rather, practitioners should should take repeated
measurements (e.g.~in MRI and borehole tomography), increase apparatus
sensitivity (e.g.~in EIT), or take consecutive measurements (e.g.~in
the 1D heat equation). Overall, we show that measurement
clusterization is a natural and (almost) inevitable part of Bayesian
D-optimal designs.

### 2024-07-21 latex/refs.tex
before diff

\section{Reviewer 1}
This paper explains a common practical issue: the cause of
clusterization in Bayesian D-optimal designs in infinite dimensions
and why adding a correlated measurement process decluster the design
points. Its theoretical results are valuable to the field of inverse
problems/uncertainty quantification and the conclusions should be
useful to practitioners in this field.  However, I feel changing the
assumption from independent to correlated measurements is quite
artificial: people tweak a data generation process assumption to solve
an optimal design problem, and the author just takes the assumption as
it is. Since the author proves the clusterization issue under the
independent measurement assumption, I expect the author to go deeper
and explain why an ordinary data generation assumption + a common
optimal design leads to a “suboptimal” design? Not just by the
mathematical proof, but from an information theory point of view. For
example, does the redundancy points mean additional points does not
reduce the variance of the estimation at all? If that’s the case, how
does correlated measurement assumption “help” with it?

@@

Nearly all textbooks on classical design of experiments (e.g. Morris,
2011, Section 1.2.4) will feature a section with the three words
randomisation, blocking and repetition: the important principles of
design of experiments.  In these texts, repeated measurements are
considered to be a beneficial property of a design. Interestingly,
when optimal designs are found, they are often found to have repeated
measurements.  The paper does not explain why the author thinks
repeated measurements (measurement clusterisation) is an undesirable
property. If one looks at, for example, Fedorov (1996) and Nyberg et
al (2012), it is an undesirable property because they are interested
in sensor location, and two or more sensors cannot be in the same
location. However, there are many experiments where repeated
measurements are possible, and are therefore optimal (if the model is
correct, which it won’t be). An example, is a chemical engineering
experiment where one wishes to learn the relationship between a series
of variables and yield of a chemical reaction. The experiment will
involve specifying the variables, and measuring the yield (the
observation) after a specified period of time. This process is then
duplicated with a potentially different set of variables.  It is
perfectly possible to have repeated variables.

@@

\item The paper explains why repeated measurements can be optimal
  (answer to Question 3). The author should expand on this and draw
  the link with how repeated measurements are, typically, regarded as
  desirable in classical design of experiments for finite
  dimensions. See for example, the discussion of the relationship
  between Caratheodory’s Theorem and the number of support points in
  Pronzato \& Pazman (2013, page 139). The support points are the
  unique design points, so if this is less than m, then there are
  repeated design points.
\end{enumerate}

>>>>

\section{Reviewer 1}
This paper explains a common practical issue: the cause of
clusterization in Bayesian D-optimal designs in infinite dimensions
and why adding a correlated measurement process decluster the design
points. Its theoretical results are valuable to the field of inverse
problems/uncertainty quantification and the conclusions should be
useful to practitioners in this field.

\answer: I appreciate the kind words.

However, I feel changing the assumption from independent to correlated
measurements is quite artificial: people tweak a data generation
process assumption to solve an optimal design problem, and the author
just takes the assumption as it is. Since the author proves the
clusterization issue under the independent measurement assumption, I
expect the author to go deeper and explain why an ordinary data
generation assumption + a common optimal design leads to a
“suboptimal” design? Not just by the mathematical proof, but from an
information theory point of view. For example, does the redundancy
points mean additional points does not reduce the variance of the
estimation at all? If that’s the case, how does correlated measurement
assumption “help” with it?

\answer: Thanks for your questions. I am not a proponent of using
correlated noise, I just wanted to show the usefullness of my model by
rigorously proving how correlated noise mitigates clusterization. I
have added several lines clearing this issue.

@@

\answer:

Nearly all textbooks on classical design of experiments (e.g. Morris,
2011, Section 1.2.4) will feature a section with the three words
randomisation, blocking and repetition: the important principles of
design of experiments.  In these texts, repeated measurements are
considered to be a beneficial property of a design. Interestingly,
when optimal designs are found, they are often found to have repeated
measurements. The paper does not explain why the author thinks
repeated measurements (measurement clusterisation) is an undesirable
property.

\answer: I do not view clusterization as an undesired property, I only
mentioned many authors consider it to be undesirable. I have added a
section contrasting repetition and replication with clusterization and
why I believe they may be different.

If one looks at, for example, Fedorov (1996) and Nyberg et
al (2012), it is an undesirable property because they are interested
in sensor location, and two or more sensors cannot be in the same
location. However, there are many experiments where repeated
measurements are possible, and are therefore optimal (if the model is
correct, which it won’t be). An example, is a chemical engineering
experiment where one wishes to learn the relationship between a series
of variables and yield of a chemical reaction. The experiment will
involve specifying the variables, and measuring the yield (the
observation) after a specified period of time. This process is then
duplicated with a potentially different set of variables.  It is
perfectly possible to have repeated variables.

\answer: again

@@

  \answer: I do consider experiments where repeated measurements are
  possible. I am giving insight on why such designs are useful, even
  though repeated measurements may be non-intuitive in certain
  settings.

\item The paper explains why repeated measurements can be optimal
  (answer to Question 3). The author should expand on this and draw
  the link with how repeated measurements are, typically, regarded as
  desirable in classical design of experiments for finite
  dimensions. See for example, the discussion of the relationship
  between Caratheodory’s Theorem and the number of support points in
  Pronzato \& Pazman (2013, page 139). The support points are the
  unique design points, so if this is less than m, then there are
  repeated design points.

  \answer: great comment
\end{enumerate}
